### 2017-03-02 thesis/discussion/discussion.tex
Started on limitations section - tmp commit

@@

\subsection{\gls{blca} data set and obesity metagenes}
\label{sub:blca_and_obesity_metagenes}

@@

\subsection{False positives in gene expression analyses}
\label{sub:false_positives_in_gene_expression_analyses}

@@

\subsection{Linear models to predict obesity metagenes}
\label{sub:linear_models_to_predict_obesity_metagenes}

@@

\subsection{Significance of the pathway associated signatures with obesity associated signatures}
\label{sub:significance_of_pr_pathway}

% -- PR may have been significant only because it was from Cris' data?

The linear models that were created from the ``consistent'' pathway signatures in Cris' data set showed significant association of the \gls{pr} pathway with the obesity metagenes.
But was this due to the true biological relationship between the \gls{pr} pathway and the obesity metagenes, or only because the models were created in Cris' data set?
In fact, the proportion of the samples that were \gls{pr}$^+$ was greater in Cris' data set (62 \gls{pr}$^+$ and 37 \gls{pr}$^-$) compared with the Creighton's data set (48 \gls{pr}$^+$ and 50 \gls{pr}$^-$), which implied that the Cris' data set was biased towards \gls{pr} status and hence \gls{pr} pathway signature.

@@

In the step wise linear models, many variables were identified to be significantly associated with the obesity metagenes.
With that said, even if these variables were related to the obesity associated genetic signatures, the fact that these results depend on the pathway associated genetic signatures from the \citet{Gatza2010a} study makes the results questionable.
This was due to the lack of consistency of the metagenes across different cancer data sets, where more than half of the pathway metagenes showed a correlation of less than 0.8 in at least one data set (\cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes,app:b}), which meant that at least half of the signatures were specific to the Gatza \textit{et al.} data set.
Thus, without further validation of the quality of the pathway associated genetic signatures and verification of the roles of these signatures in the biological relationship between obesity and cancer, these results should be interpreted with great caution.

@@

\subsection{Definition of obesity}
\label{sub:definition_of_obesity}

@@

@@

@@

There were two main aims to this project: firstly to determine whether there were any obesity specific genetic signatures that could be tranferred across multiple cancer types; and secondly to investigate whether there were any biological pathways dysregulated in the samples that were obese compared to the samples that were not obese.
These aims were addressed in \cref{cha:obesity_genetic_signatures_and_cancer,cha:obesity_associated_genetic_signature_and_pathway_signatures}, resepectively.
\\

\noindent

>>>>

% There were two main aims to this project: firstly to determine whether there were any obesity specific genetic signatures that could be tranferred across multiple cancer types; and secondly to investigate whether there were any biological pathways dysregulated in the samples that were obese compared to the samples that were not obese.
% These aims were addressed in \cref{cha:obesity_genetic_signatures_and_cancer,cha:obesity_associated_genetic_signature_and_pathway_signatures}, resepectively, and will be answered in \cref{sec:conclusion}.

@@

\subsection{\gls{blca} data set and the obesity metagenes}
\label{sub:blca_and_obesity_metagenes}

@@

\subsection{False positives in the gene expression analyses}
\label{sub:false_positives_in_gene_expression_analyses}

@@

\subsection{Linear models to predict the obesity metagenes}
\label{sub:linear_models_to_predict_obesity_metagenes}

@@

\subsection{Significance of the pathway associated signatures with the obesity associated signatures}
\label{sub:significance_of_pr_pathway}

The linear models that were created from the ``consistent'' pathway signatures in Cris' data set showed significant association of the \gls{pr} pathway with the obesity metagenes.
But was this due to the true biological relationship between the \gls{pr} pathway and the obesity metagenes, or only because the models were created in Cris' data set?
In fact, the proportion of the samples that were \gls{pr}$^+$ was greater in Cris' data set (62 \gls{pr}$^+$ and 37 \gls{pr}$^-$ samples) compared with the Creighton's data set (48 \gls{pr}$^+$ and 50 \gls{pr}$^-$ samples), which implied that the Cris' data set was biased towards \gls{pr} status and hence \gls{pr} pathway signature.

@@

In the step wise linear models, many variables were identified to be significantly associated with the obesity metagenes.
With that said, even if these variables were related to the obesity associated genetic signatures, the fact that these results depend on the pathway associated genetic signatures from the \citet{Gatza2010a} study makes the results questionable.
This was due to the lack of consistency of the metagenes across different cancer data sets, where more than half of the pathway metagenes showed a correlation of less than 0.8 in at least one data set (\cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes,app:b}).
This meant that at least half of the signatures were specific to the Gatza \textit{et al.} data set and could be highly variable in other data sets.
Thus, without further validation of the quality of the pathway associated genetic signatures and verification of the roles of these signatures in the biological relationship between obesity and cancer, these results should be interpreted with great caution.

@@

% There were many limitations to this project, though some have already been mentioned in \cref{sec:summary_of_the_results}.

\subsection{Definition of obesity}
\label{sub:discussion_definition_of_obesity}

First and possibly the most important limitation to this project was the use of \gls{bmi} as a measurement for obesity and its subsequent classification of the samples based on \gls{bmi}.
It was important to define the obese phenotype accurately, as all of the obesity associated genetic signatures, gene expression analyses and pathway enrichment analysis were dependent on the patient \gls{bmi} status.
\Gls{bmi} is perhaps the most popular measurement to estimate the central adiposity owing to the relative ease of measurement in clinical settings, and hence used in various clinical trials and large epidemiological studies as an indication of the patients' metabolic state.
However, there have been studies that showed other measurements, such as waist-to-height ratio and waist-to-hip ratio, to be a better representation of the central adiposity of the patient \citep{Dalton2003,Lee2008}.

Perhaps the reason why there were so many false positives when gene expression analyses were carried out in \cref{sub:false_positives_in_gene_expression_analyses} was because \gls{bmi} was used to classify the patients into their corresponding \gls{bmi} statuses, instead of another measurement.
It was likely that the sample \gls{bmi} (and therefore \gls{bmi} status) was not good enough a phenotypic characteristic to describe the underlying biology driven by obesity.
Since the gene expression analyses were based on such ``vague'' parameter, it may have picked up so many genes that were apparently related to obesity when in fact the genes were not related to obesity at all.

Ideally, multiple measurements that are indicative of obesity should be made in order to determine the obesity status of the patients to make a more accurate classification.
% Taking these extra measurements could be the difference between 
Additional measurements such as blood lipid profiles and insulin levels may also help understand the underlying biological state in which the patients are in, and would greatly help focus on the genes that are truly affecting the patients.
However, these measurements will be difficult to carry out in a systematic fashion in a clinical setting, and to collate existing patient data with these measurements to make a data set that is large enough for the analyses done in this project may be impossible.
Nevertheless, even the measurement of waist circumference will significantly help describe the obesity phenotype more accurately in future investigations.

@@

Second limitation was the quality of the data sets in which this project have used.
All of the data sets used in this project were from an online and publicly available source, and the types of data sets used were either microarray or \gls{rnaseq} data.

First problem -- different studies generated different data sets in a different way/method

second problem -- micorarry vs. rna seq data

third problem -- pure quality of the data

@@

The final limitation to consider was the quality of the genetic signatures used in this project.
This limitation is partly related to the previous section, as all of the genetic signatures were derived from the data sets used by other groups.

@@

There were two main aims to this project: firstly to determine whether there were any obesity specific genetic signatures that could be tranferred across multiple cancer types; and secondly to investigate whether there were any biological pathways dysregulated in the samples that were obese compared to the samples that were not obese.
These aims were addressed in \cref{cha:obesity_genetic_signatures_and_cancer,cha:obesity_associated_genetic_signature_and_pathway_signatures}, resepectively.

As shown clearly by the results from \cref{cha:obesity_genetic_signatures_and_cancer}, there seemed to be no  genetic signature that were able to differntiate between the samples that were obese from those that were not, across different cancer types.

Furthermore, there seemed to be no genes that were differentially expressed between the obese and the non-obese patients that were also common between different cancer types.
